---
title: "Joins"
format: html
editor: visual
---

```{r}
library(tidyverse)
library(palmerpenguins)
library(lubridate)
library(kableExtra)
```

# Joins

```{r}
animals <- tibble::tribble(
  ~location,   ~species,  ~maturity,
   "lagoon",   "bobcat",    "adult",
    "bluff",   "coyote", "juvenile",
    "creek",      "fox",    "adult",
     "oaks", "squirrel", "juvenile",
    "bluff",   "bobcat",    "adult"
  )

```

```{r}
sites <- tibble::tribble(
  ~location,      ~full_site_name, ~jurisdiction,
   "beach",       "Goleta Beach",     "SB City",
  "lagoon",        "UCSB Lagoon",        "UCSB",
   "bluff",       "Ellwood Mesa",     "SB City",
    "oaks", "Fremont Campground",        "USFS"
  )

```

## Full Join

```{r}
full_join_example <- full_join(animals, sites)
# KEY --> by "x" = "y" 
```

## Left Join

```{r}
left_join_example <- left_join(animals, sites, by = "location")

left_join_example
```

## Inner Join

```{r}
inner_join_example <- inner_join(animals,sites, by = 'location')
inner_join_example
```

```{r}

# sites_2 <- sites |> 
#   rename((loc = location))
# 
# sites_2
```

## Semi Join

```{r}
semi_join_example <- semi_join(animals, sites, by = 'location')

semi_join_example
```

## Anti Join

```{r}
anti_join_example <- anti_join(animals, sites, by = "location")
anti_join_example
```

Refresher 1: Starting with the penguins dataset in the palmerpenguins package, write a single piped sequence in which you:

Only include penguins at Biscoe and Dream islands Remove the year and sex variables Add a new column called body_mass_kg, with penguin mass converted from grams to kilograms Rename the island variable to location

```{r}
refresher1 <- penguins |> 
  filter(island == c("Biscoe", "Dream islands")) |> 
  select(-year, -sex) |> 
  mutate(body_mass_kg = body_mass_g /1000) |> 
  rename(location = island)

names(refresher1)
head(refresher1)
```

Refresher 2: Staring with the penguins dataset in the palmerpenguins package, write a single piped sequence in which you:

Limit to only Adelie penguins Remove any observations where flipper_length_mm is NA (hint: !is.na()) Group the data by sex Find the mean (mean()), standard deviation (sd) and sample size (n) of flipper_length_mm for male and female Adelie penguins, returned in a nice summary table

1996-02-18T20:34:00-07:00

2002-12-05T02:41:00-07:00

# Dates and times with `lubridate`

```{r}
my_date <- "03-15-1998"
class(my_date)
# mdy, ymd, dmy
lubridate::mdy(my_date)

my_date2 <- "11/7/2003"
#dmy
lubridate::dmy(my_date2)

my_date3 <- "08-Jun-1974"
lubridate::dmy(my_date3)

my_date4 <- "19610618"
my_real_date <- lubridate::ymd(my_date4)
```

It can fail when using the wrong function

```{r}
lubridate::ymd("1942- 08-30")

lubridate::dmy("09/22/84")
```

```{r}
time <- "2020-08-12 11:18"
time <- lubridate::ymd_hm(time)
class(time)
```

# convert time zones

```{r}

with_tz(time, "Asia/Hong_Kong")

my_time <- lubridate::ymd_hms(time)

year(my_time)
month(my_time)
hour(my_time)
minute(my_time)
second(my_time)

# Useful for group by months
# use with mutate
# can isolate months and do analysis 
#Can ask questions such as what day or week
```

```{r}
teddy_bday <- lubridate::ymd(20170615)
wday(teddy_bday, label = TRUE)

day_born <- lubridate::ymd(19840812)
wday(day_born, label = TRUE)

week(teddy_bday)
```

```{r}
urchin_counts <- tribble(
  ~date, ~species, ~size_mm,
  "10/3/2020", 'purple', 55,
  "10/4/2020", 'red', 48,
  "11/17/2020", "red", 67
)

urchin_counts

#overwrite existing date as ISO 8601
urchin_ymd <- urchin_counts |> 
  mutate(date = mdy(date)) |> 
  mutate(year = year(date),
         month = month(date),
         day = day(date))
urchin_ymd

```

```{r}
day_1 <- ymd("2020-01-06")
day_2 <- ymd("2020-05-18")

time_interval <- interval(day_1,day_2)
time_interval

time_length(time_interval, "year") # returns units in years can be seconds --> class = numeric
```

# stringr for working with strings

## str_detect

```{r}
my_string <- "Teddy loves eating salmon and socks."

my_string |>  str_detect("love")

my_string |> 
  str_detect("pup")

my_string  <- c("burrito", "fish taco", "taco salad")

my_string |> 
  str_detect("fish")
```

```{r}
starwars

# use str_detect to detect true or false

skywalkers <- starwars |> 
  filter(str_detect(name, pattern = "Skywalker"))

skywalkers

mauls <- starwars |> 
  filter(str_detect(name, pattern = "Maul"))

mauls

blue_eyes <- starwars |> 
  filter(str_detect(eye_color, pattern = "blue"))
blue_eyes
```

# str_replace to replace string patterns

```{r}
firewalkers <- starwars |> 
  mutate(name_2 = str_replace(name, pattern = "Sky",replacement = "Fire"))

firewalkers
```

# Removing excess white space

```{r}
feedback <- c("  I ate some   nachos ", "Wednesday morning  ")
#str_trim removes extra white space on ends

str_trim(feedback)

# str_squish removes both ends and inner extra white space
str_squish(feedback)
```

# Letter Case

```{r}
str_to_lower(feedback)
str_to_upper(feedback)
str_to_title(feedback)
str_to_sentence(feedback)


# From star wars, create a new data frame starwars_lc that adds a new column called name_lower containing character names converted to all lowercase

starwars_lc <- starwars |> 
  mutate(name_lower = str_to_lower(name))

starwars_lc
```

# `case_when()`\<- alt if-else statements

```{r}
sw_robot <- starwars |> 
  mutate(robot_status = case_when(
    species == "Droid" ~ "A ROBOT!!!",
    species == "Human" ~ "It's a human!!!",
    TRUE ~ "This is something else"
  ))

sw_robot
```

# Tidiverts
